Formalizing Stack Safety as a Security Property

### 2023-02-04 submission5/core.tex
done with 1,2 pass

In the next section, we give a brief overview of our framework and
assumptions. In \cref{sec:example}, we walk through a function call in
a simple example machine and discuss how each of our properties
applies to it, informally. In the process we motivate the properties
from a security perspective.
%
In \cref{sec:formal} we formalize the machine model,
its {\em security semantics}, and the stack safety properties built on these.
In \Cref{sec:enforcement} we describe the micro-policies that we test,
in \cref{sec:testing} the testing framework itself, and in
\cref{sec:relwork,sec:future} related and future work.

@@

\apt{Still working from here down}

Exactly which operations and rules are needed depends on
what code features we wish to support.
The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
\cref{tab:psi}.\bcp{I was confused by the table because I looked at
  the figure before I read that sentence!} Those that we explore in detail in \cref{sec:example,sec:formal}
are shown in white: calls, returns, and the allocation and deallocation of local memory.
In \cref{sec:extensions}, we cover additional operations and and parameters
to support various forms of memory sharing \ifexceptions , exceptions, \fi
and tail-call elimination; these are shown in light gray. These are also
the operations we test in \cref{sec:testing}.
We give a proof-of-concept formalization of a more sophisticated, heap-like sharing model
based on pointer provenance in \cref{app:ptr}\bcp{first
  mention of this appendix; if it's important, mention it in the
  outline in section I.  Otherwise drop it.}, represented by the dark gray
operations, but we do not test this.

@@

@@

\begin{table}
\begin{center}
  \begin{tabular}{| l | l |}
    \hline
    \(\psi \in \Psi\)\bcp{huh?} & Parameters \\
    \hline
    \example \(\mathbf{call}\) & target address, argument registers \\
    \testing & stack arguments (base, offset \& size) \\
    \example \(\mathbf{return}\) & \\
    \example \(\mathbf{alloc}\) & offset \& size \\
    \testing & public flag \\
    \example \(\mathbf{dealloc}\) & offset \& size \\
    \testing \(\mathbf{tail call}\) & (same as for \(\mathbf{call}\)) \\
    \hline
    \multicolumn{2}{|c|}{{\it Pointer provenance operations}} \\
    \hline
    \theory \(\mathbf{promote}\) & register, offset \& size \\
    \theory \(\mathbf{propagate}\) & source register/address \\
    \theory & destination register/address \\
    \theory \(\mathbf{clear}\) & target register/address \\
    \hline
  \end{tabular}
\end{center}
\caption{Security-relevant operations and their parameters.}
  \label{tab:psi}
\end{table}

@@

State elements that are outside of the stack---general-purpose memory used for
globals and the heap, as well as the code region and globally-relevant
registers\bcp{what is a globally-relevant
register?}---are always labeled \(\public\). We place security requirements on some
\(\public\) elements for purposes of our \(\wbcf\) property, and a
given enforcement mechanism
might restrict their access (e.g., by rendering code immutable), but for integrity
and confidentiality purposes they are considered accessible at all times.

@@

Arguments are marked \(\object\), meaning that their contents may be safely used.
When the active function\bcp{are functions active, or are their
  arguments active?} allocates memory, that memory will also be \(\object\).
Then, on a call, \(\object\) elements that are not being used to communicate with
the new callee will become \(\sealed\)---reserved for an inactive principal,
and expected to be unchanged when it becomes active again.

@@

When our properties are used to evaluate a system, the threat model
will depend on the details of that system. However, there are some
constraints that our design puts on any system. In particular, we must
trust that the method of distinguishing security-relevant operations
is accurate.  For example, if the enforcement mnechanism involves
labels placed on code by a compiler, we must trust that the compiler
placed those labels correctly. For instance, if a compiled function
call is not marked as such, then the caller's data might not be
protected from the callee.  Conversely, marking too many operations as
calls will simply cause otherwise safe programs to be rejected.

>>>>

In the next section, we give a brief overview of our framework and
assumptions. In \cref{sec:example}, we walk through a function call in
a simple example machine and discuss how each of our properties
applies to it, informally. In the process we motivate the properties
from a security perspective.
%
In \cref{sec:formal} we formalize the machine model,
its {\em security semantics}, and the stack safety properties built on these.
In \Cref{sec:enforcement} we describe the micro-policies that we test,
in \cref{sec:testing} the testing framework itself, and in
\cref{sec:relwork,sec:future} related and future work.
\Cref{app:ptr} gives an (untested) proof-of-concept formalization of a more sophisticated,
heap-like sharing model based on pointer provenance.

@@

Exactly which operations and rules are needed depends on
what code features we wish to support.
The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
\cref{tab:psi}. A core set of operations covering calls, returns, and local
memory is introduced in the example in \cref{sec:example}
and formalized in \cref{sec:formal}. An extended set covering simple memory sharing and
tail-call elimination are described in \cref{sec:extensions} and tested in \cref{sec:testing}.
The remaining operations are needed for the capability-based model in \cref{app:ptr}. 
%\bcp{I was confused by the table because I looked at
% the figure before I read that sentence!}
%Those that we explore in detail in \cref{sec:example,sec:formal}
%are shown in white: calls, returns, and the allocation and deallocation of local memory.
%In \cref{sec:extensions}, we cover additional operations and and parameters
%to support various forms of memory sharing \ifexceptions , exceptions, \fi
%and tail-call elimination; these are shown in light gray. These are also
%the operations we test in \cref{sec:testing}.

@@

\apt{Hopefully better now.}

@@

\begin{table}
\begin{center}
  \begin{tabular}{| l | l | l |}
    \hline
    Operation \(\psi \in \Psi\) & Parameters & Sections\\
    \hline
    \example \(\mathbf{call}\) & target address, argument registers & \ref{sec:example},\ref{sec:formal}\\
    \testing & stack arguments (base, offset \& size) & \ref{sec:extensions},\ref{sec:testing} \\
    \example \(\mathbf{return}\) & & \ref{sec:example},\ref{sec:formal}\\
    \example \(\mathbf{alloc}\) & offset \& size & \ref{sec:example},\ref{sec:formal}\\
    \testing & public flag & \ref{sec:extensions},\ref{sec:testing} \\
    \example \(\mathbf{dealloc}\) & offset \& size &  \ref{sec:example},\ref{sec:formal}\\
    \testing \(\mathbf{tail call}\) & (same as for \(\mathbf{call}\)) & \ref{sec:extensions},\ref{sec:testing} \\
%    \hline
%    \multicolumn{2}{|c|}{{\it Pointer provenance operations}} \\
%    \hline
    \theory \(\mathbf{promote}\) & register, offset \& size & \ref{app:ptr} \\
    \theory \(\mathbf{propagate}\) & source register/address & \ref{app:ptr} \\
    \theory & destination register/address & \ref{app:ptr} \\
    \theory \(\mathbf{clear}\) & target register/address & \ref{app:ptr} \\
    \hline
  \end{tabular}
\end{center}
\caption{Security-relevant operations and their parameters, with the Sections in which
  they are first used in the paper.}
  \label{tab:psi}
\end{table}

@@

State elements that are outside of the stack---general-purpose memory used for
globals and the heap, as well as the code region and globally shared
registers---are always labeled \(\public\). We place security requirements on some
\(\public\) elements for purposes of the \(\wbcf\) property, and a
given enforcement mechanism
might restrict their access (e.g., by rendering code immutable), but for integrity
and confidentiality purposes they are considered accessible at all times.

@@

Arguments are marked \(\object\), meaning that their contents may be safely used.
When a function allocates memory for its own stack frame, that memory will also be \(\object\).
Then, on a call, \(\object\) elements that are not being used to communicate with
the new callee will become \(\sealed\)---reserved for an inactive principal,
and expected to be unchanged when it becomes active again.

@@

When our properties are used to evaluate a system, the threat model
will depend on the details of that system. However, there are some
constraints that our design puts on any system. In particular, we must
trust that the security-relevant operations have been correctly labeled.
If a compiled function
call is not marked as such, then the caller's data might not be
protected from the callee; conversely, marking too many operations as
calls will simply cause otherwise safe programs to be rejected.
